Very simplified likelihood

### 2023-11-23 inst/very_simplified_likelihood.Rmd
more work

f_single_direct <- function(a,r,n,log=FALSE){  # not vectorised in anything
  b <- (1-a)/n   # 0 <= b <= 1/n
  if(log){
      return(log(1-n*b)+r*log(b)-sum(log(1-seq_len(r)*b)))
     } else {
      return((1-n*b)*b^r-prod(1-(1:3)*b))
     }
}

f <- function(a,r,n,log=FALSE){
out <- a*0
if(log){
    for(rr in r){  out <- out + f_single(a,rr,n,log=TRUE)}
  } else {
    out[] <- 1
    for(rr in r){  out <- out * f_single(a,rr,n,log=FALSE)}
  }
  return(out)
}

@@

```{r}
 a <- seq(from=0.01,to=0.99,by=0.01)
 L <- f(a,r=c(3,4,5),10)
plot(a,L-max(L))
```

>>>>

f_single_direct <- function(a,r,n,log=FALSE){  # not vectorised in anything
  b <- (1-a)/n   # 0 <= b <= 1/n
  if(log){
      return(log(a) + r*log(b) - sum(log(1-seq_len(r)*b)))
     } else {
      return(a * b^r/prod(1-seq_len(r)*b))
     }
}

`f_vectorised_direct` <- function(a,r,n,log=FALSE){  # vectorised in 'a', not in 'r'
  b <- (1-a)/n   # 0 <= b <= 1/n
  out <- r*log(b) + log(a) -rowSums(log(1-outer(b,seq_len(r))))
  if(log){return(out)} else {return(exp(out))}
}

f <- function(a,r,n,log=FALSE){
out <- a*0
if(log){
    for(rr in r){  out <- out + f_vectorised_direct(a,rr,n,log=TRUE)}
  } else {
    out[] <- 1
    for(rr in r){  out <- out * f_vectorised_direct(a,rr,n,log=FALSE)}
  }
  return(out)
}

@@

```{r}
 a <- seq(from=0.05,to=0.45,by=0.01)
 L <- f(a,r=c(3,4,3,3,1,3,3,3,1,1,1,5),10,log=T)  # 10+1==11 competitors in total
plot(a,L-max(L))
abline(v=a[which.max(L)])
abline(v=1/11)
abline(h=c(0,-2))
```

Now skating:

```{r showhughes}
skating_table
hughes <- unlist(skating_table[1,])
a <- seq(from=0.05,to=0.55,by=0.01)
L <- f(a,hughes,23,log=TRUE)
L <- L - max(L)
plot(a,L,type='b')
abline(v=1/23)
```

```{r showall}
a <- seq(from=0.001,to=0.55,by=0.001)
fun <- function(i){
  out <- f(a,unlist(skating_table[i,]),23,log=TRUE)
  return(out-max(out))
}

plot(NA,xlim=c(0,1),ylim=c(-10,0),type='n')
abline(v=1/23)
for(i in seq_len(9)){points(a,fun(i),type='b')}

plot(NA,xlim=c(-5,1),ylim=c(-5,0),type='n',xlab="log-odds")
abline(v=-log(22))
for(i in seq_len(23)){points(log(a/(1-a)),fun(i),type='l')}
```

# Formula 1

```{r showformula1}
x <- read.table("formula1_2022.txt",header=TRUE)
x <- x[,-ncol(x)]
x
```

```{r analyseform1}
fun <- function(i){
  suppressWarnings(jj <- as.numeric(unlist(x[i,])))
  jj[is.na(jj)] <- 23
  L <- f(a,jj,23,log=TRUE)
  return(L - max(L))
}

plot(NA,xlim=c(-5,1),ylim=c(-5,1),type='n',xlab="log-odds")
abline(v=-log(22))
for(i in seq_len(20)){points(log(a/(1-a)),fun(i),type='l')}

```

# Mathematics Olympiad

```{r}
x <- as.matrix(read.table("olympiad.txt",header=TRUE))
x <- x[apply(x,1,max) <= 100,]
fun <- function(i){
 jj <- as.numeric(unlist(x[i,]))
  L <- f(a,jj,100,log=TRUE)
  return(L - max(L))
}
plot(NA,xlim=c(-5,1),ylim=c(-5,1),type='n',xlab="log-odds")
abline(v=-log(99))
for(i in seq_len(40)){points(log(a/(1-a)),fun(i),type='l')}

```{r}
x <- as.matrix(read.table("ol.txt",header=TRUE))[,1:4]
x <- x[apply(x,1,max) <= 100,]
fun <- function(i){
 jj <- as.numeric(unlist(x[i,]))
  L <- f(a,jj,100,log=TRUE)
  return(L - max(L))
}
plot(NA,xlim=c(-5,1),ylim=c(-5,1),type='n',xlab="log-odds")
abline(v=-log(99))
for(i in seq_len(40)){points(log(a/(1-a)),fun(i),type='l')}
```

```{r}
x <- as.matrix(read.table("ol.txt",header=TRUE))[,1:4]
x <- x[apply(x,1,max) <= 100,]
x[] <- sample(c(x))
x
rowSums(x)
fun <- function(i){
 jj <- as.numeric(unlist(x[i,]))
  L <- f(a,jj,100,log=TRUE)
  return(L - max(L))
}
plot(NA,xlim=c(-5,1),ylim=c(-5,1),type='n',xlab="log-odds")
abline(v=-log(99))
for(i in seq_len(40)){points(log(a/(1-a)),fun(i),type='l')}
